---
title: "InstaCart Predictive Data Prep: Predictive Data Prep Alternative Samples"
author: "April Leclair, Daniel Ochoa, Hoang Anh Thai Vu, Qiuhan Sun"
date: "December 11, 2017"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "PredictiveAnalysis_HTML_Output") })
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
---

```{r setup, include=FALSE, cache=FALSE}
library(tidyverse)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```



# Prepare the Data

## Load the Data

```{r message=FALSE, warning=FALSE, cache=TRUE}
load("../Source/data_final.Rda")
load("../Source/order_products_sample60_prior.Rda")
products <- read_csv("../Source/products.csv")  
```




# Top 1000 Users and Products

## Identify the Top 1000 Users

```{r}
topUsers_1000 <- order_products_sample60_prior %>%
  left_join(products, by="product_id") %>%
  mutate(user_id = as.factor(user_id)) %>%
  group_by(user_id) %>%
  summarise(uid.numProducts=n()) %>%
  arrange(desc(uid.numProducts)) %>%
  head(n=1000)
```

## Identify the Top 1000 Most Popular Products

```{r}
topProducts_1000 <- order_products_sample60_prior %>%
  left_join(products, by="product_id") %>%
  mutate(product_id = as.factor(product_id)) %>%
  group_by(product_id, product_name) %>%
  summarise(product.numProducts=n()) %>%
  arrange(desc(product.numProducts)) %>%
  head(n=1000)
```

## Sample Dataset with the Top 1000 Users and Products

```{r}
df.uid1000 <- subset(data_final, user_id %in% topUsers_1000$user_id) 
df.uid1000 <- subset(df.uid1000, product_id %in% topProducts_1000$product_id) 
df.uid1000 <- as.data.frame(df.uid1000)     
```

## Save the Final Sampled Dataset

```{r}
# save(df.uid1000, file = "../Source/df.uid1000.Rda")
```

## Clean

```{r}
rm(df.uid1000, products, topProducts_1000, topUsers_1000)
```






# Random 1000 Users who Ordered at Least 10 Times and More than 50 Products 

## Identify Users who Have Ordered More Than 100 Products

```{r}
uid.numProducts_100g <- data_final %>%
  group_by(user_id) %>%
  summarise(uid.numProducts=mean(uid.numProducts)) %>%
  filter(uid.numProducts >= 100) %>%
  arrange(desc(uid.numProducts)) 
```

## Identify Users who Have Made More Than 15 Orders

```{r}
uid.lastOrder_15g <- order_products_sample60_prior %>% 
  group_by(user_id) %>% 
  summarise(uid.lastOrder = max(order_number)) %>%
  filter(uid.lastOrder >= 15) %>%
  arrange(desc(uid.lastOrder)) 
```

# Sample Data Set with Users who Have Made More Than 15 Orders & More Than 100 Purchase of Products

```{r}
df.ordProd <- 
  subset(data_final, user_id %in% uid.lastOrder_15g$user_id) 
df.ordProd <- 
  subset(df.ordProd, user_id %in% uid.numProducts_100g$user_id) 

length(unique(df.ordProd$user_id)) # 11509 users in the data table
```

# Still Big, so Sample 900 & 500 Random Users

## 900
```{r}
set.seed(888)  
random_users <- sample(unique(df.ordProd$user_id), 900)  
df.ordProd900 <-
  subset(df.ordProd, user_id %in% random_users)
```

## 500
```{r}
set.seed(555)  
random_users <- sample(unique(df.ordProd$user_id), 500)  
df.ordProd500 <-
  subset(df.ordProd, user_id %in% random_users)
```

## As Data Frame
```{r}
df.ordProd900 <- as.data.frame(df.ordProd900)       
df.ordProd500 <- as.data.frame(df.ordProd500)   
```

## Save the Final Sampled Dataset

```{r}
# save(df.ordProd900, file = "../Source/df.ordProd900.Rda")
# save(df.ordProd500, file = "../Source/df.ordProd500.Rda")
```

# Clean 

```{r}
rm(data_final, df.ordProd, df.ordProd500, df.ordProd900, order_products_sample60_prior, uid.lastOrder_15g, uid.numProducts_100g, random_users)
```