---
title: "InstaCart Predictive Analysis: Number of Items (Danny)"
author: "April Leclair, Daniel Ochoa, Hoang Anh Thai Vu, Qiuhan Sun"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "PredictiveAnalysis_HTML_Output") })
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
---

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(tree)
library(randomForest)
```

This file is used to predict how many items a user will order on their next order

# Reading In The Data
```{r message=FALSE, warning=FALSE }
orders <- read_csv("../Source/orders_sample40.csv") %>%
  mutate(order_dow = as.character(recode(order_dow, `0`="Sunday", `1`="Monday", `2`="Tuesday",
                            `3`="Wednesday", `4`="Thurday", `5`="Friday", `6`="Saturday")))
order_products <- read_csv("../Source/order_products_sample40.csv")
```

# Preparing the Data

Below we create a num_products variable for the total number of products ordered in each order.
```{r}
orders <-
  order_products %>%
  group_by(order_id) %>%
  summarise(num_products = n()) %>%
  right_join(orders, by="order_id")
```

Below we create a avg_num_products variable for the average number of products this user orders.
```{r}
orders <-
  orders %>%
  group_by(user_id) %>%
  mutate(avg_num_products = mean(num_products))
```

Below we create a previous_num_products variable for the number of products in this users previous order.
```{r}
orders <-
  orders %>%
  group_by(user_id) %>%
  mutate(previous_num_products = dplyr::lag(num_products, order_by=user_id))
```


Once we have sampled our training and test datasets we can move on to making predictions.
```{r echo=FALSE}
ordersTrain <-
  orders %>%
  filter(eval_set == "train") %>%
  select(user_id, order_id, order_number, num_products, eval_set, previous_num_products, avg_num_products, order_dow, order_hour_of_day, days_since_prior_order)
ordersPrior <- 
  orders %>%
  filter(eval_set == "prior") %>%
  select(user_id, order_id, order_number, num_products, eval_set, previous_num_products, avg_num_products, order_dow, order_hour_of_day, days_since_prior_order)
```

# Making Predictions

## Predictions using a Regression Tree
```{r echo=FALSE}
regTree <- tree(num_products ~ days_since_prior_order+avg_num_products+previous_num_products, ordersPrior)
summary(regTree)
plot(regTree)
text(regTree, pretty=1)
```

## Predictions using a Random Forest

I have to lower ntrees (taking too long, look up other optimaizations)
```{r}
Rforest <- randomForest(num_products ~ days_since_prior_order+avg_num_products+previous_num_products, ordersTrain, na.action=na.roughfix)
summary(Rforest)
```


