---
title: "InstaCart Predictive Analysis: Hierarchical Clustering and K-Means by `user_id`"
author: "April Leclair, Daniel Ochoa, Hoang Anh Thai Vu, Qiuhan Sun"
date: "November 25, 2017"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "PredictiveAnalysis_HTML_Output") })
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
---

```{r setup, include=FALSE, cache=FALSE}
library(tufte)
library(readr)
library(tidyverse)
library(ggplot2)
library(ape)                #to make hierarchical clusters
library(mclust)             #for k-means clustering
library(BBmisc)             #for data normalization
library(dplyr)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```



# Load the data

```{r message=FALSE, warning=FALSE, cache=FALSE}
orders <- read_csv("../Source/orders_sample.csv")
order_products <- read_csv("../Source/order_products_sample.csv")
products <- read_csv("../Source/products.csv")
departments <- read_csv("../Source/departments.csv")
aisles <- read_csv("../Source/aisles.csv")
```

```{r}
names(orders)
names(order_products)
names(products)
names(departments)
names(aisles)
```





# Prepare the data: by `user_id` without `order_chron`

## Filter users who ordered more than 50 products

Note: The `user_id` column is replaced as a rowname(df) in all data tables

```{r}
uid_train_nochron <- orders %>%
  left_join(order_products, by="order_id") %>%
  left_join(products, by="product_id") %>%
  filter(eval_set == "prior") %>%
  mutate(user_id = as.factor(user_id),
         days_since_prior_order = ifelse(is.na(days_since_prior_order), 0, days_since_prior_order)) %>%
  select(user_id, order_number, product_name, reordered, days_since_prior_order,
         order_dow, order_hour_of_day, add_to_cart_order, order_id) %>%
  arrange(user_id, order_number)

uid_train_nochron <- uid_train_nochron %>%
  group_by(user_id) %>%
  summarize(num_products=n(), # total number of products bought; repetition not overwritten
            reorder_products=sum(reordered),
            reord_ord_ratio=reorder_products/num_products,
            days_since=mean(days_since_prior_order),
            min_days_since=min(days_since_prior_order),
            max_days_since=max(days_since_prior_order),
            sd_days_since=sd(days_since_prior_order),
            ave_hr=mean(as.numeric(order_hour_of_day)),
            sd_hr=sd(as.numeric(order_hour_of_day)),
            ave_dow=mean(as.numeric(as.factor(order_dow)) ,
            sd_dow=sd(as.numeric(as.factor(order_dow))))) %>%
  filter(num_products >= 50) 
temp <- uid_train_nochron
uid_train_nochron <- as.data.frame(uid_train_nochron)
rownames(uid_train_nochron) <- uid_train_nochron$user_id
uid_train_nochron[,1] <- NULL

head(uid_train_nochron)
glimpse(uid_train_nochron)         # 38,996
anyNA(uid_train_nochron)           # FALSE
```

## Filter users who ordered more than 100 products

```{r}
uid_train_nochron_100 <- temp %>%
  filter(num_products >= 100)
uid_train_nochron_100 <- as.data.frame(uid_train_nochron_100)
rownames(uid_train_nochron_100) <- uid_train_nochron_100$user_id
uid_train_nochron_100[,1] <- NULL
save(uid_train_nochron_100, file="../Source/uid_priorsamp_nochron_100.Rda")

head(uid_train_nochron_100)
glimpse(uid_train_nochron_100)     # 17,512
anyNA(uid_train_nochron_100)       # FALSE
```

## Filter users who ordered more than 200 products

```{r}
uid_train_nochron_200 <- temp %>%
  filter(num_products >= 200)
uid_train_nochron_200 <- as.data.frame(uid_train_nochron_200)
rownames(uid_train_nochron_200) <- uid_train_nochron_200$user_id
uid_train_nochron_200[,1] <- NULL
save(uid_train_nochron_200, file="../Source/uid_priorsamp_nochron_200.Rda")

head(uid_train_nochron_200)
glimpse(uid_train_nochron_200)     # 9,315
anyNA(uid_train_nochron_200)       # FALSE
```





# Inspect the data table `uid_train_nochron_200`

## Take a glimpse of `uid_train_nochron_200`

```{r}
glimpse(uid_train_nochron_200)
```

## Check if any variable has missing values

```{r}
sum(is.na(uid_train_nochron_200$num_products))
sum(is.na(uid_train_nochron_200$reorder_products))
sum(is.na(uid_train_nochron_200$days_since))
```

## Examine some key statistics of the variables

### `num_products` across `user_id`

```{r}
min(uid_train_nochron_200$num_products)
max(uid_train_nochron_200$num_products)
mean(uid_train_nochron_200$num_products)
sd(uid_train_nochron_200$num_products)
# hist(uid_train_nochron_200$num_products, main=paste("Histogram of", colnames(uid_train_nochron_200)[1]))
```

### `reorder_products` across `user_id`

```{r}
min(uid_train_nochron_200$reorder_products)
max(uid_train_nochron_200$reorder_products)
mean(uid_train_nochron_200$reorder_products)
sd(uid_train_nochron_200$reorder_products)
# hist(uid_train_nochron_200$reorder_products, main=paste("Histogram of", colnames(uid_train_nochron_200)[2]))
```

### `reord_ord_ratio` across `user_id`

```{r}
mean(uid_train_nochron_200$reord_ord_ratio)
sd(uid_train_nochron_200$reord_ord_ratio)
```

### `days_since` across `user_id`

```{r}
min(uid_train_nochron_200$days_since)
max(uid_train_nochron_200$days_since)
mean(uid_train_nochron_200$days_since)
sd(uid_train_nochron_200$days_since)
```

### `min_days_since` across `user_id`

```{r}
min(uid_train_nochron_200$min_days_since)
max(uid_train_nochron_200$min_days_since)
mean(uid_train_nochron_200$min_days_since)
sd(uid_train_nochron_200$min_days_since)
```

### `max_days_since` across `user_id`

```{r}
min(uid_train_nochron_200$max_days_since)
max(uid_train_nochron_200$max_days_since)
mean(uid_train_nochron_200$max_days_since)
sd(uid_train_nochron_200$max_days_since)
```

### `sd_days_since` across `user_id`

```{r}
min(uid_train_nochron_200$sd_days_since)
max(uid_train_nochron_200$sd_days_since)
mean(uid_train_nochron_200$sd_days_since)
sd(uid_train_nochron_200$sd_days_since)
```





# Process the data for clustering

## Sample, normalize, and compute similarity with the data set of users who ordered more than 200 products

```{r}
uid_train_nochron_200_samp <- uid_train_nochron_200[sample(1:nrow(uid_train_nochron_200), 1000, replace=FALSE),]
order_norm_200_samp <- normalize(uid_train_nochron_200_samp)
sapply(order_norm_200_samp, mean)
sapply(order_norm_200_samp, var)
set.seed(15)
order_norm_200_samp_diffs <- dist(order_norm_200_samp)
anyNA(order_norm_200_samp_diffs)
```

## Keep the unsampled data set of users who ordered more than 200 products

```{r}
order_norm_200 <- normalize(uid_train_nochron_200)
sapply(order_norm_200, mean)
sapply(order_norm_200, var)
set.seed(10)
order_norm_200_diffs <- dist(order_norm_200)
anyNA(order_norm_200_diffs)
```

## Plot a Dendrogram of 1000 random users who ordered more than 200 products

```{r,fig.width=7,fig.height=100}
order_norm_200_samp_diffs %>%
  hclust("ave") %>%
  as.phylo() %>%
  plot(cex=0.7, label.offset=0, main="Dendrogram of 1000 Random Users with No Chronology", sub="who ordered more than 200 products")
```

## Plot a Dendrogram of all users who ordered more than 200 products

### Create a closeup data set for an enhanced visual representation

```{r}
order_norm_200_diffs_closeup <- order_norm_200_diffs %>%
  hclust("ave")
d1 <- cut(as.dendrogram(order_norm_200_diffs_closeup), h=1)
d10 <- cut(as.dendrogram(order_norm_200_diffs_closeup), h=10)
```

### Plot some branches of the closeup data set

```{r}
# $lower[[129]]
# 'dendrogram' with 2 branches and 2 members total, at height 0.9451381 
plot(d1$lower[[129]])

# $lower[[1]]
# 'dendrogram' with 2 branches and 3 members total, at height 5.014784 
plot(d10$lower[[1]])
```





# Process the data for **k-Means** clustering

## Perform k-means clustering on normalized data `order_norm_200_samp`:

```{r}
set.seed(111)
num_clusters <- 4
kmean <- order_norm_200_samp
order_clusts <- kmean %>%
  kmeans(centers = num_clusters)%>%
  fitted("classes")%>%
  as.character()
kmean <- kmean %>%
  mutate(cluster=order_clusts)
```

```{r,echo=FALSE}
ggplot(kmean, aes(x=days_since, y=reorder_products))+
  geom_point(aes(color=cluster),alpha=.25)
```

### `cluster` == 1 from `kmean`

```{r,echo=FALSE}
ggplot(kmean,aes(x=days_since,y=reorder_products))+
  geom_point(aes(color=cluster, alpha=cluster==1))  
```

### `cluster` == 2 from `kmean`

```{r,echo=FALSE}
ggplot(kmean,aes(x=days_since,y=reorder_products))+
  geom_point(aes(color=cluster, alpha=cluster==2))    
```

### `cluster` == 3 from `kmean`

```{r,echo=FALSE}
ggplot(kmean,aes(x=days_since,y=reorder_products))+
  geom_point(aes(color=cluster, alpha=cluster==3))    
```

### `cluster` == 4 from `kmean`

```{r,echo=FALSE}
ggplot(kmean,aes(x=days_since,y=reorder_products))+
  geom_point(aes(color=cluster, alpha=cluster==4))    
```
