---
title: "InstaCart Predictive Analysis: Logit Using `df.ordProd`"
author: "April Leclair, Daniel Ochoa, Hoang Anh Thai Vu, Qiuhan Sun"
date: "December 11, 2017"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "PredictiveAnalysis_HTML_Output") })
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
---

```{r setup, include=FALSE, cache=FALSE}
library(tufte)
library(caret)  # Used for confusion matrix
library(tidyverse)
library(pROC)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```





# Prepare the Data

## Load the Data

```{r message=FALSE, warning=FALSE, cache=FALSE}
load("../Source/df.ordProd.Rda")
```

```{r}
df.ordProd <- df.ordProd %>% 
  mutate(user_id = as.numeric(user_id),
         product_id = as.numeric(product_id))
```


## Divide Train & Test Data

```{r cache.lazy=FALSE}
samplesize <- floor(0.70 * nrow(df.ordProd))
set.seed(098)
all_indices = c(1:nrow(df.ordProd))
train_indices = sample(seq_len(nrow(df.ordProd)), size = samplesize)
test_indices = setdiff(all_indices, train_indices)
all( ! test_indices %in% train_indices)

# Train_data
train = df.ordProd[train_indices, ]
train <- train %>% arrange(user_id, product_id)

# Test_data
test = df.ordProd[test_indices, ] 
test <- test %>% arrange(user_id, product_id)

rm(samplesize, all_indices, train_indices, test_indices, df.ordProd)
```

## Expand Memory Storage

```{r}
memory.limit()
memory.limit(size=60000)
```





# Predictive Analysis: Logistic Regression

## Full Model 

Fit the model for binomial regression.
```{r}
modelf <- glm(reordered ~ . - `user_id` - `product_id`, family = binomial(link = 'logit'), data = train)
```

Obtain the results of our model:
```{r}
summary(modelf)
```


## Fewer Variables

Fit the model for binomial regression. We remove these statistically insignificant variables this time.

`uid.reorderProducts`, `uid.maxDaysSince`, `uid.aveDow`, `uid.sdDow`, `uid.pantry_distr` & `uid.bakery_distr`
```{r}
model2 <- glm(reordered ~ . - `user_id` - `product_id` - `uid.numProducts` - `uid.reorderProducts` -  `uid.sdDaysSince` -  `uid.accDaysSince` - `uid.aveHr`  - `uid.sdHr` - `uid.aveDow` - `uid.sdDow` - `uid.frozen_distr` - `uid.bakery_distr` - `uid.deli_distr`, family = binomial(link = 'logit'), data = train)
```

Obtain the results of our model:
```{r}
summary(model2)
```


## Even Fewer Variables

Remove variables with one star from `model2`

`uid.snacks_distr` & `uid.frozen_distr`
```{r}
model3 <- glm(reordered ~ . - `user_id` - `product_id` - `uid.numProducts` - `uid.reorderProducts` -  `uid.sdDaysSince` -  `uid.accDaysSince` - `uid.aveHr`  - `uid.sdHr` - `uid.aveDow` - `uid.sdDow` - `uid.frozen_distr` - `uid.bakery_distr` - `uid.deli_distr` -`uid.maxDaysSince` - `uid.pantry_distr`, family = binomial(link = 'logit'), data = train)
```

Obtain the results of our model:
```{r}
summary(model3)
```


## Compare Models

```{r}
anova(modelf, model2, test="Chisq")
anova(modelf, model3, test="Chisq")
anova(model2, model3, test="Chisq")
```

It means that the fitted model "modelf" is not significantly different from either "model2" or "model3". However, "model3" is significantly different form "model2" at p = 0.1. 


## Testing Function

```{r}
f1_test <- function (pred, ref,user_id) {
  require(ModelMetrics)
  dt <- tibble(user_id, pred, ref)
  dt <- dt %>%
    group_by(user_id)%>%
    mutate(f1_score = f1Score(pred,ref))%>%
    summarise(f1_score = mean(f1_score,na.rm=TRUE))
  f1_mean <- mean(dt$f1_score,na.rm=TRUE)
  return (f1_mean)
}
```


## Null Model 

```{r}
nullPredict <- ifelse(test$user_product.order_streak > 0, 1, 0)
f1_test(nullPredict, test$reordered, test$user_id)
```


## F1 Score

Threshold not maximizer of F1, but good enough to take into account specificity.

```{r}
pred_logit_pre <- predict(modelf, newdata = test, type = 'response')
pred_logit <- ifelse(pred_logit_pre > 0.53, 1, 0)
pred_logit[is.na(pred_logit)]<-0
pred_f1 <- f1_test(pred_logit, test$reordered, test$user_id)
pred_f1
```


## Plot GLM

```{r warning = FALSE}
pred_logit_pre <- predict(model3, newdata = test, type = 'response')

ggplot(test, aes(x=pred_logit_pre, y=reordered)) + geom_point() +
  stat_smooth(method="glm", family="binomial", se=TRUE) +
  labs(x="Prediction", y="Actual",
       title="Logistic Regression of Prediction vs. Actual Reordered",
       caption="Data from InstaCart Kaggle Competition")
```


## ROC

Threshold close to being a maximizer for AUC.

```{r warning = FALSE}
pred_logit <- ifelse(pred_logit_pre > 0.10, 1, 0)
roc(test$reordered, pred_logit, plot=TRUE)
```