---
title: "InstaCart Exploratory Analysis"
author: "April Leclair, Daniel Ochoa, Hoang Anh Thai Vu, Qiuhan Sun"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "ExploratoryAnalysis_HTML_Output") })
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
---

This file provides some exploratory analysis of the InstaCart data.

```{r setup, include=FALSE}
library(tufte)
library(tidyverse)
library(ggplot2)
library(scales)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Reading in the data
```{r cache=TRUE}
orders = read.csv("../Source/orders.csv")
```

Let's start off looking at some rows in the order data.
```{r}
head(orders, 12)
```

Quick Overview of the data: Hello

order_id is a unique value given to each order.  
user_id is a unique value given to each user.  
eval_set refers the set of data that this row belongs to. This attribute has three levels: "prior", "train", and "test". The model should be trained on "prior", validated on "train" and ran on "test" for submission.  
order_number refers to the nth order that this this user had made.  
order_dow refers to the day of the week, 0 being Sunday (not confirmed), that the order was made.  
order_hour_of_day refers to the hour of the day that the order was made.  
days_since_prior_order refers to the numnber of days that have passed since this user last made an instaCart order.  

We can see from the 12 rows above that the rows are ordered according to user_id and within users is ordered according order_number. 

Let's start trying to understand the data:

# When Do People Order?

```{r warning=FALSE}
ggplot(data=orders, aes(x=order_hour_of_day)) + geom_histogram(stat="count", fill="seagreen4") + scale_y_continuous(labels = comma) + labs(x="Hour of day", y ="Number of orders", title="Orders by Hour of Day")
```
It appears that the majority of orders are made between 10am and 4pm.