---
title: "InstaCart Exploratory Analysis: Users - 153479 & 44294"
author: "April Leclair, Daniel Ochoa, Hoang Anh Thai Vu, Qiuhan Sun"
date: "November 25, 2017"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "ExploratoryAnalysis_HTML_Output") })
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
---

```{r setup, include=FALSE, cache=FALSE}
library(tufte)
library(readr)
library(tidyverse)
library(ggplot2)
library(ape)                #to make hierarchical clusters
library(mclust)             #for k-means clustering
library(BBmisc)             #for data normalization
library(dplyr)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```


This document compares two users that were in the same category from the hierarchcal clustering analysis. We look at users with `user_id` == 153479 & 44294.


# Load the data

```{r message=FALSE, warning=FALSE, cache=FALSE}
orders <- read_csv("../Source/orders_sample.csv")
order_products <- read_csv("../Source/order_products_sample.csv")
products <- read_csv("../Source/products.csv")
departments <- read_csv("../Source/departments.csv")
aisles <- read_csv("../Source/aisles.csv")
```

```{r}
names(orders)
names(order_products)
names(products)
names(departments)
names(aisles)
```




# Compare the buying patterns of 1000 random users who bought more than 2000 products: `user_id` == 153479 & 44294

## Prepare the data sets for each user

```{r}
user153479 <- 
  orders %>%
  filter(user_id == 153479,
         eval_set == "prior") %>% 
  left_join(order_products, by="order_id") %>%
  left_join(products, by="product_id")  
  
user44294 <- 
  orders %>%
  filter(user_id == 44294,
         eval_set == "prior") %>%  
  left_join(order_products, by="order_id") %>%
  left_join(products, by="product_id")  
```

## Conduct feature engineering

### Create a varable for discrete factored time of the day

```{r}
# new_levels <- c("Late Evening","Late Evening","Late Evening","Dawn","Dawn","Dawn","Dawn",
#                 "Morning","Morning","Morning","Morning","Morning",
#                 "Afternoon","Afternoon","Afternoon","Afternoon","Afternoon",
#                 "Evening","Evening","Evening","Evening",
#                 "Late Evening","Late Evening","Late Evening","Late Evening")
# user153479$timeofday <- factor(new_levels[user153479$order_hour_of_day])
# user153479$timeofday <- factor(user153479$order_hour_of_day, levels = c("Dawn", "Morning", "Afternoon", "Evening", "Late Evening"))
# user44294$timeofday <- factor(new_levels[user44294$order_hour_of_day])
# user44294$timeofday <- factor(user44294$timeofday, levels = c("Dawn", "Morning", "Afternoon", "Evening", "Late Evening"))
# rm(new_levels)
```

### Create a varable for continuous numeric time of the day

```{r}
user153479$timeofday_num <- as.numeric(user153479$order_hour_of_day)
user44294$timeofday_num <- as.numeric(user44294$order_hour_of_day)
```

### Create a varable for ordered factor day of the week

```{r}
levels(user153479$order_dow) <- c("Sunday", "Monday", "Tuesday", "Wednesday",
                                                 "Thursday", "Friday", "Saturday")
levels(user44294$order_dow) <- c("Sunday", "Monday", "Tuesday", "Wednesday",
                                                 "Thursday", "Friday", "Saturday")
```

## Visualize discrete day of the week

These users do not look that similar just by looking at the `days_since_prior_order` by `day of the week` plot. However, the plots reflect order chronology that the cluster data did not take into account.

```{r, warning=FALSE, fig.width=10, fig.height=7.5}
is.na(user153479$days_since_prior_order) <- 0
user153479 %>%
  na.omit() %>%
  ggplot(aes(x=order_number, y=days_since_prior_order)) +
  geom_line(color="#FF9999") +
  geom_point(aes(size = 0.5, alpha = 0.25, color=order_dow)) +
  scale_color_brewer(palette = "RdPu") +
  guides(size=FALSE, alpha=FALSE) +
  geom_smooth(stat="smooth", position="stack", method="loess") +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(title="Days Since Prior Order for User ID 153479",
       subtitle="by Time of the Day",
       color="Time of the Day",
       size="Size",
       x="Order Chronology", y="Days Since Prior Order",
       caption="Data from InstaCart Kaggle Competition")
```

```{r, warning=FALSE, fig.width=10, fig.height=7.5}
user44294 %>%
  na.omit() %>%
  ggplot(aes(x=order_number, y=days_since_prior_order)) +
  geom_line(color="#FF9999") +
  geom_point(aes(size = 0.5, alpha = 0.25, color=order_dow)) +
  scale_color_brewer(palette = "RdPu") +
  guides(size=FALSE, alpha=FALSE) +
  geom_smooth(stat="smooth", position="stack", method="loess") +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(title="Days Since Prior Order for User ID 44294",
       subtitle="by Time of the Day",
       color="Time of the Day",
       size="Size",
       x="Order Chronology", y="Days Since Prior Order",
       caption="Data from InstaCart Kaggle Competition")
```

## Table the comparisons

Tables generally reflect more similarities in the buying patterns of these users.

### Conclusion: sort of similar in `order_dow`

```{r}
# user153479: Most orders were made on Monday (65) and Tuesday (51).  
table(user153479$order_dow) 

# user44294: Most orders were made on Monday (64) and Tuesday (64).
table(user44294$order_dow) 
```

### Conclusion: similar in `timeofday`

```{r}
# # user153479: Most orders were made in the afternoon (139) as we saw above.
# table(user153479$timeofday) 
# 
# # user44294: Most orders were made in the afternoon (163) as we saw above.
# table(user44294$timeofday) 
```

### Conclusion: similar in `days_since_prior_order`

```{r}
# user153479: Ten days (47) was the most common duration till the next order and the next common were four, three, twenty-one days, respectively (23, 21, 20). 
table(user153479$days_since_prior_order) 

# user44294: Eleven days (30) was the most common duration till the next order and the next common were six, thirty, seven, and eight days, respectively (28, 24, 21, 19). 
table(user44294$days_since_prior_order) 
```

### Conclusion: not so similar in `avg_days_since` by `order_dow`

```{r}
user153479 %>% 
  group_by(order_dow) %>% 
  na.omit() %>%
  summarize(avg_days_since=mean(days_since_prior_order)) 
user44294 %>% 
  group_by(order_dow) %>% 
  na.omit() %>%
  summarize(avg_days_since=mean(days_since_prior_order)) 
```

### Conclusion: not so similar in `avg_days_since` by `order_hour_of_day`

```{r}
user153479 %>% 
  group_by(order_hour_of_day) %>% 
  na.omit() %>%
  summarize(avg_days_since=mean(days_since_prior_order)) %>%
  plot()
user44294 %>% 
  group_by(order_hour_of_day) %>% 
  na.omit() %>%
  summarize(avg_days_since=mean(days_since_prior_order)) %>%
  plot()
```

### Conclusion: sort of similar in `avg_hr_of_day` and `sd_hr_of_day`

```{r}
# user153479: Most orders were made between noon and 2pm with the exception of the only order on Friday (5pm). The SD on each day of the week range from 0 to 4.  
user153479 %>% 
  group_by(order_dow) %>% 
  na.omit() %>%
  summarize(avg_hr_of_day=mean(as.numeric(order_hour_of_day)),
            sd_hr_of_day=sd(as.numeric(order_hour_of_day))) %>% head(n=10) 
# user44294: Most orders were made between noon and 4pm with the exception of the only order on Saturday (11pm). The SD on each day of the week range from 0 to 2.  
user44294 %>% 
  group_by(order_dow) %>% 
  na.omit() %>%
  summarize(avg_hr_of_day=mean(as.numeric(order_hour_of_day)),
            sd_hr_of_day=sd(as.numeric(order_hour_of_day))) %>% head(n=10)
```

## Create a more detailed visualization: Histogram of `order_dow` for `user_id` 153479 & 44294

user153479 made most orders on Mondays and Tuesdays in the the Afternoon (12-4pm), Morning (9am), and the Evening (5-6pm) in this order. The order hours are more spread out.

```{r, warning=FALSE, fig.width=10, fig.height=7.5}
user153479 %>%
  na.omit() %>%
  group_by(order_dow, order_hour_of_day) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=order_dow, y=count, fill=order_hour_of_day)) +
  geom_bar(stat="identity") +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(title="Day of the Week: Orders for User ID 153479",
       subtitle="",
       x="Day of the Week",
       y="Count",
       fill="Order Hour of Day",
       caption="Data from InstaCart Kaggle Competition")
```

user44294 made most orders on Mondays and Wednesdays in the Afternoon (2pm-4pm). The order hours are not as spread out.

```{r, warning=FALSE, fig.width=10, fig.height=7.5}
user44294 %>%
  na.omit() %>%
  group_by(order_dow, order_hour_of_day) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=order_dow, y=count, fill=order_hour_of_day)) +
  geom_bar(stat="identity") +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(title="Day of the Week: Orders for User ID 44294",
       subtitle="",
       x="Day of the Week",
       y="Count",
       fill="Order Hour of Day",
       caption="Data from InstaCart Kaggle Competition")
```
