---
title: "InstaCart Exploratory Analysis: Days Since Prior Order"
author: "April Leclair, Daniel Ochoa, Hoang Anh Thai Vu, Qiuhan Sun"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "ExploratoryAnalysis_HTML_Output") })
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
---

```{r setup, include=FALSE, cache=FALSE}
library(tufte)
library(tidyverse)
library(dplyr)
library(ggrepel)
library(lubridate)
library(stats)
library(scales)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

This file provides some exploratory analysis regarding the days_since_prior_order variable of the InstaCart data.

We know from careful consideration of the order_dow and days_since_prior_order variables that the data for each user are from consecutive orders.

# Reading In The Data

```{r cache=TRUE}
orders <- read_csv("../Source/orders.csv") %>%
  mutate(order_dow = as.character(recode(order_dow, `0`="Sunday", `1`="Monday", `2`="Tuesday",
                            `3`="Wednesday", `4`="Thurday", `5`="Friday", `6`="Saturday")))
order_products <- read_csv("../Source/order_products__prior.csv")
products <- read_csv("../Source/products.csv")
```

# Preparing The Data 

## Randomly sampling the data

We randomly sample the data because performing computations on the full dataset is expensive.

Before sampling:
  orders - 3.4 million rows  
  order_products - 32.4 million rows
  
After sampling: 293 thousand rows

```{r}
set.seed(42)  # Allows for reproducable sample results
random_users <- sample(unique(orders$user_id), 20000)
orders <-
  subset(orders, user_id %in% random_users)
```


## Removing undesired rows

Because we are primarily concerned with the `days_since_prior_order` variable we only want to consider those observations that have a value in that column. Furthermore, we only want to consider observations that are part of the `prior` evaluation set since that is the set which we will be building our model on.
```{r}
orders <- orders %>%
  filter(!is.na(days_since_prior_order)) %>%
  filter(eval_set == "prior") %>%
  group_by(user_id) %>%
  mutate(cumulative_days = cumsum(days_since_prior_order)) %>%  # This is going to come in useful later
  ungroup()
```


# How Long Until Users Reorder?
```{r warning=FALSE}
orders %>%
  ggplot(aes(x=days_since_prior_order)) + geom_histogram(stat="count", fill="seagreen4") + labs(title="Distribution of days_since_prior_order", x="Days Since Prior Order", y="Count") + scale_y_continuous(labels = comma)
```
From the visualization above we can see that most often users tend to order from InstaCart once every week (7 days). Interestingly, in second place is once a month (30 days). At first this it perplexing since the distribution of counts leading up to 30 is, in general, decreasing. However, if we assume that all of the times that users wait > 30 days to reorder is coded as 30 in the dataset than this makes a lot more sense.


# What is the Average Number of Days Since Prior Order?
```{r}
orders %>%
  filter(!is.na(days_since_prior_order)) %>%
  summarise(avg = mean(days_since_prior_order))
```

It seems that on average, an InstaCart user will go approximately 11 days  between orders.

# How does time since last order affect the rate of reordering

## Preparing the data

We need to join our orders table with the order_products table to have access to the reordered variable.
```{r cache=TRUE}
orders_products <- 
  orders %>%
  dplyr::left_join(order_products, by="order_id")  # tidyverse left_join causes discrepancy between console and knit
```

Now we construct a visualization
```{r warning=FALSE}
orders_products %>%
  group_by(days_since_prior_order) %>%
  summarise(avg_proportion_reordered = mean(reordered)) %>%
  ggplot(aes(x=days_since_prior_order, y=avg_proportion_reordered)) + geom_col(fill="seagreen4") + labs(title="How The Number Of Days Since Prior Order affects Proportion Of Items Reordered", x="Days Since Prior Order", y="Average Proportion of Items Reordered")
```
As we can see from the plot above, in general, as the number of days since prior order increases the average proportion of items that are reorders decreases. This makes sense. If a users waits 30 days before using InstaCart again perhaps that is because the items they ordered last time were a dissapointment. They aren't likely to reorder those items again.

# On Average, How Long Does It Take Each Product To Be Reordered?

## Preparing the data

We are going to create a variable `days_since_product_last_ordered` across each user and take the average of that, `avg_days_since_product_last_ordered` across all observations.
```{r}
orders_products_daysSinceProductLastOrdered <-
  orders_products %>%
  group_by(user_id) %>%
  mutate(days_since_product_last_ordered = stats::ave(cumulative_days, product_id, FUN=function(x) c(0, diff(x)))) %>%
  ungroup() %>%
  filter(!days_since_product_last_ordered == 0) %>%  # Filter out first time buying item
  group_by(product_id) %>%
  summarise(avg_days_since_product_last_ordered = mean(days_since_product_last_ordered)) %>%
  inner_join(products, by="product_id") %>%  # For product names
  arrange(avg_days_since_product_last_ordered)
```

Since we want to display the product names and we don't want a cluttered visualization we only consider the top and bottom 20 products.
```{r}
highest20 <- tail(orders_products_daysSinceProductLastOrdered, 20)
lowest20 <- head(orders_products_daysSinceProductLastOrdered, 20)
```

TODO: Requirement on number of orders? Too many 1's for top 20/100
```{r}
ggplot(data=lowest20, aes(x=product_id, y=avg_days_since_product_last_ordered)) + geom_point(alpha=0.05) + geom_text_repel(aes(label=product_name), color="seagreen4") + labs(title="The 20 Products With The Lowest Average Days Since Last Order", x="Product ID", y="Average Days Since Product Was Last Ordered")

ggplot(data=highest20, aes(x=product_id, y=avg_days_since_product_last_ordered)) + geom_point(alpha=0.05) + geom_text_repel(aes(label=product_name), color="seagreen4") + labs(title="The 20 Products With The Highest Average Days Since Last Order", x="Product ID", y="Average Days Since Product Was Last Ordered")
```

We can see that the items that, on average, go the fewest days in between orders are characteriszed by:
Candies, Ice Creams, Fruits, Tea, and Cheese.

By comparison the items that, on average, go the fewest days in between orders are characteriszed by:
Utility items: Coffee Filter Squares, Whitening Toothpaste, Seasoning, Designer Bowls, and Carpet Fresheners, as well as pills.

This makes sense, most of the items that go the longest in between orders tend to last a long time, most don't have an expiration date. Whereas the items that go the shortest in between orders tend to expire quickly or be consumed quickly.